---
title: "GIS Final"
author: "Darian"
date: "2023-12-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Load libraries
library(readr)
library(dplyr)
library(vegan)  # for diversity function
library(tidyr)
library(tidyverse)
library(ggplot2)
library(purrr)
library(car)
library(geosphere)

#Import Data

fauna_data <- read_csv("C:/Users/bradd/OneDrive - University of the Virgin Islands/UVI/Fall 2023/GIS/Final/Fauna_MetaData.csv")
View(fauna_data)

raft_data <- read_csv("C:/Users/bradd/OneDrive - University of the Virgin Islands/UVI/Fall 2023/GIS/Final/Filled_Raft_MetaData.csv")
View(raft_data)

```
# shoreline_data <- read_csv("C:/Users/bradd/OneDrive - University of the Virgin Islands/UVI/Fall 2023/GIS/Final/Shoreline_Data.csv")
```{r Calculating Raft Volume}
raft_data$Volume <- raft_data$Length * raft_data$Width * raft_data$Thickness
```

```{r Calculating Biodiversity}

# Aggregate species counts per raft
species_counts <- fauna_data %>%
  group_by(Site, Species) %>%
  summarize(Count = n(), .groups = 'drop')

# Spread the data for Shannon-Weiner Index calculation
species_matrix <- species_counts %>%
  pivot_wider(names_from = Species, values_from = Count, values_fill = list(Count = 0))

# Calculate Shannon-Weiner Index for each raft
raft_biodiversity <- species_matrix %>%
  rowwise() %>%
  mutate(Shannon_Index = diversity(c_across(-Site), index = "shannon"))

# Select only Site and Shannon_Index
raft_biodiversity <- raft_biodiversity %>%
  select(Site, Shannon_Index)

# Merge raft volume data
raft_bio <- merge(raft_data, raft_biodiversity, by = "Site")
```

# ```{r Calculating Distance From Shore}
# 
# # Assuming raft_data contains raft coordinates and a Beach
# 
# # Function to calculate the nearest distance from a raft to its beach's shoreline
# calculate_nearest_distance <- function(raft_lat, raft_lon, Beach, shoreline_data) {
#   raft_point <- cbind(raft_lon, raft_lat)
#   shore_points <- subset(shoreline_data, Beach == Beach)
#   shore_matrix <- cbind(shore_points$Longitude, shore_points$Latitude)
#   
#   # Calculate distances to all shore points for the beach and find the minimum
#   distances <- distHaversine(raft_point, shore_matrix)
#   min_distance <- min(distances)
#   return(min_distance)
# }
# 
# # Apply the function to each row in raft_data
# raft_data <- raft_data %>%
#   rowwise() %>%
#   mutate(Distance_To_Shore = calculate_nearest_distance(Latitude, Longitude, Beach, shoreline_data))
# 
# # Now raft_data includes Distance_To_Shore for each raft
# ```

# ```{r Calculate Water Quality Index}
# 
# ```

```{r Hypothesis 1 Correlation: Raft Volume vs. Biodiversity }

# Correlation analysis between Volume and Shannon Index
cor.test(raft_bio$Volume, raft_bio$Shannon_Index)

# Save the result
correlation_result <- cor.test(raft_bio$Volume, raft_bio$Shannon_Index)

# Print the results
correlation_result
```

```{r Hypothesis 1 ANOVA: Raft Type vs. Biodiversity - normally distributed}

# Checking assumptions for ANOVA
# Normality check
normal_check <- ggplot(raft_bio, aes(x = Raft_Type, y = Shannon_Index)) + 
  geom_boxplot() +
  geom_jitter(width = 0.2) +
  theme_minimal()

#Print the Results
normal_check

# Homogeneity of variances check
bartlett_anova <- bartlett.test(Shannon_Index ~ Raft_Type, data = raft_bio)

# Print the results
bartlett_anova

# Perform ANOVA or Kruskal-Wallis Test
# ANOVA
anova_result <- aov(Shannon_Index ~ Raft_Type, data = raft_bio)

anova_summary <- summary(anova_result)

anova_summary

# If ANOVA assumptions are not met, use Kruskal-Wallis test

# Post-hoc analysis if ANOVA is significant
# (Use if ANOVA result is significant and you want to know which groups differ)
# if (summary(anova_result)$`Pr(>F)`[1] < 0.05) {
#   TukeyHSD(anova_result)
# }
```

```{r Hypothesis 1 Kruskal-Wallis: Raft Type vs. Biodiversity - non-normally distributed}
kruskal_result <- kruskal.test(Shannon_Index ~ Raft_Type, data = raft_bio)
kruskal_result

```

```{r Hypothesis 1 nMDS: Community Composistion}

# Prepare a community data matrix
# Assuming fauna_data has columns 'Site' for rafts and 'Species' for species names
community_matrix <- fauna_data %>%
  group_by(Site, Species) %>%
  summarize(Count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Species, values_from = Count, values_fill = list(Count = 0))

# Remove the Site column for nMDS analysis
community_matrix <- select(community_matrix, -Site)

# Run nMDS Analysis
nmds_results <- metaMDS(community_matrix, distance = "bray", trymax = 100)

# Plot the nMDS results
plot(nmds_results$points[,1], nmds_results$points[,2], 
     xlab="NMDS1", ylab="NMDS2", main="nMDS Plot of Macrofauna Communities")
text(nmds_results$points[,1], nmds_results$points[,2], 
     labels = rownames(nmds_results$points), cex = 0.8, pos = 3)

```

# ```{r Hypothesis 2 ANOVA Assumptions: Raft Volume vs. Beaufort Force}
# 
# # Assuming 'raft_bio' is your data frame with BF and Volume
# # QQ Plot for Biodiversity Index
# BF_Vol_norm <- ggplot(raft_bio, aes(sample = Volume)) + 
#   geom_qq() + 
#   geom_qq_line() +
#   facet_wrap(~ BF) +
#   theme_minimal() +
#   ggtitle("QQ Plot for Raft Volume by Beaufort Scale")
# 
# BF_Vol_norm
# 
# # Group by Beaufort Scale and apply Shapiro-Wilk test to each group
# normality_results <- raft_bio %>%
#   group_by(BF) %>%
#   summarize(shapiro_p_value = shapiro.test(Volume)$p.value)
# 
# # View the results
# normality_results
# 
# # Levene's Test for Volume Index
# leveneTest(Shannon_Index ~ BF, data = raft_bio)
# 
# # Alternatively, Bartlett's Test
# bartlett.test(Volume ~ BF, data = raft_bio)
# ```

```{ Hypothesis 2 ANOVA vs. KW: Beaufort Force vs. Volume}

# ANOVA or Kruskal-Wallis for Raft Volume
anova_volume <- aov(Volume ~ BF, data = raft_data)
summary(anova_volume)

# If non-normal data, use Kruskal-Wallis
kruskal_volume <- kruskal.test(Volume ~ BF, data = raft_data)
kruskal_volume
```


```{r Hypothesis 2 Chi-squared: Raft Type vs. Beaufort Force}
# For raft type, a Chi-squared test might be more appropriate because Raft_Type is a categorical variable

# Assuming Raft_Type is a factor in raft_data
chisq_test <- chisq.test(table(raft_data$Raft_Type, raft_data$BF))
chisq_test
```

```{r Hypothesis 2 ANOVA Assumptions: Biodiversity vs. Beaufort Force}

# Assuming 'raft_bio' is your data frame with BF and Biodiversity
# Regression Analysis
regression_biodiversity <- lm(Shannon_Index ~ BF, data = raft_bio)
summary(regression_biodiversity)
```

```{r Hypothesis 2 Muliple Regression: Biodiversity vs. Raft Volume vs. Beaufort Force}

# Multiple Regression Analysis
multiple_regression_model <- lm(Shannon_Index ~ Volume + BF, data = raft_bio)
summary(multiple_regression_model)
```

# ```{r Hypothesis 3 Regression: Biodiversity vs. Distance from Shore}
# # Assuming 'raft_bio' is your data frame that now includes 'Distance_To_Shore' for each raft and 'Shannon_Index' as the biodiversity measure
# 
# # Model biodiversity as a function of distance from shore
# regression_model_h3 <- lm(Shannon_Index ~ Distance_To_Shore, data = raft_bio)
# 
# # View the summary of the regression model
# summary(regression_model_h3)
# 
# ```

# ```{r Hypothesis 4 Regression: Water Quality vs. Biodiversity }
# 
# #Need to come up with a water quality index
# # Assuming 'raft_bio' contains 'Water_Quality_Index' and 'Shannon_Index'
# 
# # Model biodiversity as a function of water quality
# regression_model_h4 <- lm(Shannon_Index ~ Water_Quality_Index, data = raft_bio)
# 
# # View the summary of the regression model
# summary(regression_model_h4)
# ```
