---
title: "GIS Final"
author: "Darian"
date: "2023-12-06"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Introduction

The biodiversity of marine ecosystems is a complex and dynamic attribute, critically influenced by various environmental factors. This research is dedicated to a comprehensive analysis of how different environmental variables, including raft volume, distance from shore, and water quality, affect the biodiversity associated with these Sargassum rafts. Through rigorous data analysis, we aim to shed light on the patterns and relationships that define these unique marine environments.

Utilizing a dataset encompassing various measurements and observations related to Sargassum rafts and the associated macrofauna, we will explore several hypotheses. Each hypothesis targets specific aspects of the interaction between environmental conditions and biodiversity. Our findings are expected to contribute to the broader understanding of marine biodiversity dynamics, particularly in the context of floating seaweed ecosystems.

---

```{r setup, include=FALSE}
# Load libraries
library(readr)
library(dplyr)
library(vegan)  # for diversity function
library(tidyr)
library(tidyverse)
library(ggplot2)
library(purrr)
library(car)
library(geosphere)

#Import Data

fauna_data <- read_csv("C:/Users/bradd/OneDrive - University of the Virgin Islands/UVI/Fall 2023/GIS/Final/Fauna_MetaData.csv")
View(fauna_data)

raft_data <- read_csv("C:/Users/bradd/OneDrive - University of the Virgin Islands/UVI/Fall 2023/GIS/Final/Filled_Raft_MetaData.csv")
View(raft_data)

```

## Data Set Up 

### Calculating Raft Volume
```{r Calculating Raft Volume}
raft_data$Volume <- raft_data$Length * raft_data$Width * raft_data$Thickness
```

### Calculating Biodiversity 
```{r Calculating Biodiversity}

# Aggregate species counts per raft
species_counts <- fauna_data %>%
  group_by(Site, Species) %>%
  summarize(Count = n(), .groups = 'drop')

# Spread the data for Shannon-Weiner Index calculation
species_matrix <- species_counts %>%
  pivot_wider(names_from = Species, values_from = Count, values_fill = list(Count = 0))

# Calculate Shannon-Weiner Index for each raft
raft_biodiversity <- species_matrix %>%
  rowwise() %>%
  mutate(Shannon_Index = diversity(c_across(-Site), index = "shannon"))

# Select only Site and Shannon_Index
raft_biodiversity <- raft_biodiversity %>%
  select(Site, Shannon_Index)

# Merge raft volume data
raft_bio <- merge(raft_data, raft_biodiversity, by = "Site")
```

## Hypothesis 1: The Relationship Between Sargassum Raft Volume/Type and Biodiversity {.tabset}

### Volume vs. Biodiversity - Correlation

#### Methods
```{r Correlation}
correlation_result <- cor.test(raft_bio$Volume, raft_bio$Shannon_Index)
```

#### Results
```{r}
correlation_result
```

### Raft Type vs. Biodiversity - Normal - ANOVA

#### Methods

- Check Assumptions

```{r}
# Normality check
normal_check <- ggplot(raft_bio, aes(x = Raft_Type, y = Shannon_Index)) + 
  geom_boxplot() +
  geom_jitter(width = 0.2) +
  theme_minimal()

#Print the Results
normal_check
```

Figure 1: QQ plot to assess the normality of biodiversity indices across different raft types. The plot facilitates evaluating the assumption of normality, a prerequisite for the ANOVA test.
```{r}
# Homogeneity of variances check
bartlett_anova <- bartlett.test(Shannon_Index ~ Raft_Type, data = raft_bio)

# Print the results
bartlett_anova
```

- If assumptions are met, run an ANOVA, if not, use Kruskal-Wallis test

```{r ANOVA}
anova_result <- aov(Shannon_Index ~ Raft_Type, data = raft_bio)

anova_summary <- summary(anova_result)

```

#### Results
```{r}
anova_summary
```

- Post-hoc analysis if ANOVA is significant
 (Use if ANOVA result is significant and you want to know which groups differ)
 if (summary(anova_result)$`Pr(>F)`[1] < 0.05) {
   TukeyHSD(anova_result)
 }

### Raft Type vs. Biodiversity - Non-Normal - Kruskal-Wallace

#### Methods 
```{r Kruskal-Wallace}
kruskal_result <- kruskal.test(Shannon_Index ~ Raft_Type, data = raft_bio)
```

#### Results 
```{r}
kruskal_result
```

### Community Composition - nMDS

#### Methods
```{r nMDS}

# Prepare a community data matrix
# Assuming fauna_data has columns 'Site' for rafts and 'Species' for species names
community_matrix <- fauna_data %>%
  group_by(Site, Species) %>%
  summarize(Count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Species, values_from = Count, values_fill = list(Count = 0))

# Remove the Site column for nMDS analysis
community_matrix <- select(community_matrix, -Site)
```

##### Results
```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide', fig.cap="Figure 2: Non-metric multidimensional scaling plot depicting the community composition of macrofauna associated with Sargassum rafts. The plot illustrates similarities and differences in community structures across various sampling sites."}
# Run nMDS Analysis
nmds_results <- metaMDS(community_matrix, distance = "bray", trymax = 100)

# Plot the nMDS results
plot(nmds_results$points[,1], nmds_results$points[,2], 
     xlab="NMDS1", ylab="NMDS2", main="nMDS Plot of Macrofauna Communities")
text(nmds_results$points[,1], nmds_results$points[,2], 
     labels = rownames(nmds_results$points), cex = 0.8, pos = 3)

```


## Hypothesis 2: The Relationship Between Sargassum Raft Volume/Type, Biodiversity, and Beauforts Force {.tabset}

### Raft Volume vs. Beaufort Force - Normal - ANOVA

#### Methods

- Check Assumptions

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide', fig.cap="Figure 3: Normality Check for Hypothesis 2. This Quantile-Quantile (QQ) plot assesses the normality of data related to the impact of wind/wave activity on Sargassum raft volume and biodiversity. The plot compares the distribution of sample data against a theoretical normal distribution, helping to validate the assumption of normality required for subsequent ANOVA or regression analyses."}

# Assuming 'raft_bio' is your data frame with BF and Volume
 # QQ Plot for Biodiversity Index
 BF_Vol_norm <- ggplot(raft_bio, aes(sample = Volume)) + 
   geom_qq() + 
   geom_qq_line() +
   facet_wrap(~ BF) +
   theme_minimal() +
   ggtitle("QQ Plot for Raft Volume by Beaufort Scale")
 
 BF_Vol_norm
 
 # Group by Beaufort Scale and apply Shapiro-Wilk test to each group
 normality_results <- raft_bio %>%
   group_by(BF) %>%
   summarize(shapiro_p_value = shapiro.test(Volume)$p.value)
 
 # View the results
 normality_results
```

```{r}
# Homogeneity of variances check
bartlett_anova2 <- bartlett.test(Volume ~ BF, data = raft_bio)

# Print the results
bartlett_anova2
```

- If assumptions are met, run an ANOVA, if not, use Kruskal-Wallis test

```{r}
anova_result2 <- aov(Volume ~ BF, data = raft_bio)

anova_summary2 <- summary(anova_result2)

```

#### Results

```{r}
anova_summary2
```

- Post-hoc analysis if ANOVA is significant
 (Use if ANOVA result is significant and you want to know which groups differ)
 if (summary(anova_result)$`Pr(>F)`[1] < 0.05) {
   TukeyHSD(anova_result)
 }
 
### Raft Volume vs. Beaufort Force - Non-Normal - Kruskal - Wallace

#### Methods 
```{r}
kruskal_result2 <- kruskal.test(Volume ~ BF, data = raft_bio)
```

#### Results 
```{r}
kruskal_result2
```

### Raft Type vs. Beaufort Force - Non-Normal - Chi-square 

#### Methods 
```{r Hypothesis 2 Chi-squared: Raft Type vs. Beaufort Force}
# For raft type, a Chi-squared test might be more appropriate because Raft_Type is a categorical variable

# Assuming Raft_Type is a factor in raft_data
chisq_test <- chisq.test(table(raft_data$Raft_Type, raft_data$BF))
```

#### Results 
```{r}
chisq_test
```

### Biodiversity vs. Beaufort Force - Regression

#### Methods

```{r}
# Assuming 'raft_bio' is your data frame with BF and Biodiversity

regression_biodiversity <- lm(Shannon_Index ~ BF, data = raft_bio)
```

#### Results 

```{r}
summary(regression_biodiversity)

```

### Biodiversity vs. Volume vs. Beaufort Force - Multiple Regression

#### Methods

```{r}
multiple_regression_model <- lm(Shannon_Index ~ Volume + BF, data = raft_bio)
```

#### Results 

```{r}
summary(multiple_regression_model)
```

